---
title: "eSports_Visualizations"
output: html_notebook
---
# Introduction {.tabset}

Chess is a classical board game that became a big e-sport game in the recent years. Specially with the pandemic we see more and more games and competitions being hosted online. Therefore, chess is considerate one of the e-sports of the moment with crescent views in streaming websites such as Twitch. In this project, utilized a data from Kaggle that contains more than 20000 thousand games played in the Lichess website. I produced some visualizations to ilustrate some of theinteresting aspects of these 20000 thousand games. I hope you enjoy it!

The data set is available at https://www.kaggle.com/datasnaek/chess.


In order to build the visualizations I utilized the package tidyverse (ggplot).

```{r}
library(tidyverse)
```


```{r}
chess <- read.csv("games.csv")
```

## Explorign the data

The data has 20,058 rows (records) and 16 columns(features).


```{r}
chess
```

```{r}
View(chess)
```


## Difference in Rating Versus the Winner (rated matches)

In this visualization, we see the winners of the online chess matches and their difference in rating. 

```{r}
winners_diff <- chess %>%
  select(rated, winner, white_rating, black_rating) %>% 
  filter(rated == "TRUE") %>%
  select(winner, white_rating, black_rating) %>% 
  mutate(rating_diff = white_rating - black_rating) %>%
  mutate(rating_diff_class = if_else(rating_diff > 100, "plus_100_eloRating_white",
                                     if_else(rating_diff < -100, "plus_100_eloRating_black", "even_elo")))
```

We selected the features winner, white_rating, black_rating. We caculated the difference betweent the ratings and assign them into classes (i.e., even_elo, plus_100_eloRating_white, and plus_100_eloRating_black).

```{r}
winners_diff
```



```{r}
ggplot(data = winners_diff, mapping = aes(x = rating_diff_class)) +
  geom_histogram(mapping = aes(fill = winner, colour = winner), stat = "count", position = "dodge") +
  geom_text(mapping = aes(fill = winner, label =..count..),
            vjust=-0.5,
            stat="count",
            position = position_dodge(width = 0.9)) +
  ylim(0, max(1800)) +
  scale_fill_grey(start = 0.2, end = 1) +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(panel.background = element_rect(fill = "cornsilk"),
        plot.background = element_rect(fill = "cornsilk3"),
        legend.background = element_rect(fill = "cornsilk3", colour = NA),
        panel.grid = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.25, color="black" ),
        plot.title = element_text(hjust = 0.5, size = 16),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)) +
  labs(title = "Number of Rated Matches Won According to Rating ♞",
       y = "Number of Wins",
       x = "Difference in elo Rating",
       caption = "Source: Kaggle")
  
```

```{r}
rating_outcome_plot <- ggplot(data = winners_diff, mapping = aes(x = rating_diff_class)) +
  geom_histogram(mapping = aes(fill = winner, colour = winner), stat = "count", position = "dodge") +
  geom_text(mapping = aes(fill = winner, label =..count..),
            vjust=-0.5,
            stat="count",
            position = position_dodge(width = 0.9)) +
  ylim(0, max(1800)) +
  scale_fill_grey(start = 0.2, end = 1) +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(panel.background = element_rect(fill = "cornsilk"),
        plot.background = element_rect(fill = "cornsilk3"),
        legend.background = element_rect(fill = "cornsilk3", colour = NA),
        panel.grid = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.25, color="black" ),
        plot.title = element_text(hjust = 0.5, size = 16),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)) +
  labs(title = "Number of Rated Matches Won According to Rating ♞",
       y = "Number of Wins",
       x = "Difference in elo Rating",
       caption = "Source: Kaggle")

ggsave(plot = rating_outcome_plot, file = "rating_outcomeplot.png", height = 8, width = 12)
```

## Manner of the Win, a comparison between even and uneven elo ratiing


```{r}
victory_status_data <- chess %>%
  select(rated, white_rating, black_rating, victory_status) %>% 
  filter(rated == "TRUE") %>%
  select(white_rating, black_rating, victory_status) %>% 
  mutate(rating_diff = white_rating - black_rating) %>%
  mutate(rating_diff_class = if_else(rating_diff > 100, "uneven_elo",
                                     if_else(rating_diff < -100, "uneven_elo", "even_elo"))) %>%
  select(victory_status, rating_diff_class)
```



```{r}
victory_status_data
```



```{r}
ggplot(data = victory_status_data, mapping = aes(x = rating_diff_class)) +
  geom_histogram(mapping = aes(fill = victory_status, colour = victory_status), stat = "count", position = "dodge") +
  geom_text(mapping = aes(fill = victory_status, label =..count..),
            vjust=-0.5,
            stat="count",
            position = position_dodge(width = 0.9)) +
  ylim(0, max(2100)) +
  scale_fill_grey(start = 0.2, end = 1) +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(panel.background = element_rect(fill = "cornsilk"),
        plot.background = element_rect(fill = "cornsilk3"),
        legend.background = element_rect(fill = "cornsilk3", colour = NA),
        panel.grid = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.25, color="black" ),
        plot.title = element_text(hjust = 0.3, size = 16),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)) +
  labs(title = "Number of Rated Matches Won According to Victory Status ♞",
       y = "Number of Wins",
       x = "Type of elo Rating",
       caption = "Source: Kaggle")
```

```{r}
rating_victory_status_plot <- ggplot(data = victory_status_data, mapping = aes(x = rating_diff_class)) +
  geom_histogram(mapping = aes(fill = victory_status, colour = victory_status), stat = "count", position = "dodge") +
  geom_text(mapping = aes(fill = victory_status, label =..count..),
            vjust=-0.5,
            stat="count",
            position = position_dodge(width = 0.9)) +
  ylim(0, max(2100)) +
  scale_fill_grey(start = 0.2, end = 1) +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(panel.background = element_rect(fill = "cornsilk"),
        plot.background = element_rect(fill = "cornsilk3"),
        legend.background = element_rect(fill = "cornsilk3", colour = NA),
        panel.grid = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.25, color="black" ),
        plot.title = element_text(hjust = 0.3, size = 16),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)) +
  labs(title = "Number of Rated Matches Won According to Victory Status ♞",
       y = "Number of Wins",
       x = "Type of elo Rating",
       caption = "Source: Kaggle")

ggsave(plot = rating_victory_status_plot,
       filename = "victory_status.png",
       height = 8,
       width =  12)
```

